DOM review: Data preparation

### 2023-11-27 Analysis/Archive/Code/1_DataPrep.Rmd
minor restructure

## Objectives 
This sheet contains information on the authors' stated research objectives and whether their papers produced certain products (maps/predictions). This sheet is relatively simple and does not require checks.
```{r Process objectives, echo = FALSE}
objectives_sheet <- spreadsheet[["Objectives and Inference"]] %>%
  # Select only completed papers
  filter(Status == "Complete") %>%
  # Convert objective columns to TRUE/FALSE
  mutate(Obj_Methods = case_when(`Methods?` == "Y" ~ TRUE,
                                 `Methods?` == "N" ~ FALSE),
         Obj_Projections = case_when(`Future predictions?` == "Y" ~ TRUE,
                                     `Future predictions?` == "N" ~ FALSE),
         Obj_Spatial = case_when(`Spatial predictions?` == "Y" ~ TRUE,
                                 `Spatial predictions?` == "N" ~ FALSE),
         Obj_Trends = case_when(`Observing trends?` == "Y" ~ TRUE,
                                `Observing trends?` == "N" ~ FALSE),
         Obj_Hypothesis = case_when(`Hypothesis testing?` == "Y" ~ TRUE,
                                    `Hypothesis testing?` == "N" ~ FALSE)) %>%
  # Drop unneeded columns
  select(`Review ID`, Title, Obj_Methods, Obj_Projections, 
         Obj_Spatial, Obj_Trends, Obj_Hypothesis)

@@

## Species and Location
This sheet describes the study species and area. Some papers with multi-species models may have multiple entries.
```{r Process Species/Location, echo = FALSE}
taxa_location_sheet <- spreadsheet[["Species and Location"]] %>%
  # Filter to included papers
  filter(Status == "Complete") %>%
  # Create binary threat columns
  mutate(Threat_HabitatLoss = str_detect(`Threat categories`, "HabitatLoss"),
         Threat_Exploitation = str_detect(`Threat categories`, "Exploitation"),
         Threat_ClimateChange = str_detect(`Threat categories`, "ClimateChange"),
         Threat_Invasives = str_detect(`Threat categories`, "Invasives"),
         Threat_Disease = str_detect(`Threat categories`, "Disease"),
         Threat_Disaster = str_detect(`Threat categories`, "Disturbance")) %>%
  # Split 'taxa keywords' column by commas
  separate_rows(`Taxa keywords`, sep = ", ") %>%
  # Split country column by commas
  separate_rows(Country, sep = ", ") %>%
  # Select only necessary fields
  select(`Review ID`, Title, Taxa, `Taxa keywords`, Number, 
         Threat_HabitatLoss, Threat_Exploitation, Threat_ClimateChange,
         Threat_Invasives, Threat_Disease, Threat_Disaster,
         Country, Continent, Realm, `Study Size`)
kable(head(taxa_location_sheet))
```

@@

```{r Process covariates, echo = FALSE}
covariates_sheet <- spreadsheet[["Covariates"]] %>%
  # Filter down to only included papers
  filter(Status == "Complete") %>%
  # Note modifiers for parameters, then split them off from parameters
  mutate(MultiSpecies = case_when(str_detect(Parameter,"MS") ~ TRUE,
                                  !str_detect(Parameter, "MS") ~ FALSE),
         FalsePositive = case_when(str_detect(Parameter, "FP") & str_detect(Parameter, "Detection") ~ TRUE)) %>%
  mutate(Parameter = str_split(Parameter, "_", simplify = TRUE)[,1]) %>%
  # Split out each covariate
  separate_rows(`Categories`, sep =  ", ") %>%
  # Convert 'none' to NA
  separate(`Categories`, c("Category", "Scale", "Dynanimity"), sep = "_") %>%
  select(`Review ID`, Title, `Total covariates`, Parameter, Category, Scale, 
         Dynanimity, `Higher order`, `Interactive terms`,
         MultiSpecies, FalsePositive)

covariates_sheet$Category[covariates_sheet$Category %in% c("None", "NA")] <- NA
covariates_sheet$Category[covariates_sheet$Category == c("Topography")] <- "Topographic"
covariates_sheet$Category[covariates_sheet$Category == c("Hydrology")] <- "Hydrologic"

@@

## Modelling
This contains information about the model selection and evaluation process
```{r Process modelling, echo = FALSE}
modelling_sheet <- spreadsheet[["Modelling"]] %>%
  filter(Status == "Complete") %>%
  separate_rows(`Selection category`, sep = ", ") %>%
  separate(`Selection category`, into = c("Select_Category", "Select_Type"), sep = "_") %>%
  select(`Review ID`, Title, Bayesian, Select_Category, Select_Type)

modelling_sheet$Select_Category[modelling_sheet$Select_Category %in% c("APriori", "Aprioiri")] <- "Apriori"

@@

```{r Save output, echo = FALSE}

reviewResults <- list(Study = study_sheet,
                      Objectives = objectives_sheet,
                      Taxa = taxa_location_sheet,
                      Survey = survey_sheet,
                      Covariates = covariates_sheet,
                      Modelling = modelling_sheet)

>>>>

## Objectives 
This sheet contains information on the authors' stated research objectives and whether their papers produced certain products (maps/predictions). This sheet is relatively simple and does not require checks.
```{r Process objectives, echo = FALSE}
objectives_sheet <- spreadsheet[["Objectives and Inference"]] %>%
  # Select only completed papers
  filter(Status == "Complete") %>%
  # Convert objective columns to TRUE/FALSE
  mutate(Obj_Methods = case_when(`Methods?` == "Y" ~ TRUE,
                                 `Methods?` == "N" ~ FALSE),
         Obj_Projections = case_when(`Future predictions?` == "Y" ~ TRUE,
                                     `Future predictions?` == "N" ~ FALSE),
         Obj_Spatial = case_when(`Spatial predictions?` == "Y" ~ TRUE,
                                 `Spatial predictions?` == "N" ~ FALSE),
         Obj_Hypothesis = case_when(`Hypothesis testing?` == "Y" ~ TRUE,
                                    `Hypothesis testing?` == "N" ~ FALSE)) %>%
  # Drop unneeded columns
  select(`Review ID`, Title, Obj_Methods, Obj_Projections, 
         Obj_Spatial, Obj_Hypothesis)

@@

## Species and Location
This sheet describes the study species and area. Some papers with multi-species models may have multiple entries.
```{r Process Species/Location, echo = FALSE}
taxa_location_sheet <- spreadsheet[["Species and Location"]] %>%
  # Filter to included papers
  filter(Status == "Complete") %>%
  # Create binary threat columns
  mutate(Threat_HabitatLoss = str_detect(`Threat categories`, "HabitatLoss"),
         Threat_Exploitation = str_detect(`Threat categories`, "Exploitation"),
         Threat_ClimateChange = str_detect(`Threat categories`, "ClimateChange"),
         Threat_Invasives = str_detect(`Threat categories`, "Invasives"),
         Threat_Disease = str_detect(`Threat categories`, "Disease"),
         Threat_Disaster = str_detect(`Threat categories`, "Disturbance")) %>%
  # Split 'taxa keywords' column by commas
  separate_rows(`Taxa keywords`, sep = ", ") %>%
  # Split country column by commas
  separate_rows(Country, sep = ", ") %>%
  # Select only necessary fields
  select(`Review ID`, Title, Taxa, `Taxa keywords`, `IUCN Status`, Number, 
         Threat_HabitatLoss, Threat_Exploitation, Threat_ClimateChange,
         Threat_Invasives, Threat_Disease, Threat_Disaster,
         Country, Continent, Realm, `Study Size`, `Center point`)
kable(head(taxa_location_sheet))
```

@@

```{r Process covariates, echo = FALSE}
covariates_sheet <- spreadsheet[["Covariates"]] %>%
  # Filter down to only included papers
  filter(Status == "Complete") %>%
  # Select parameter prefix
  mutate(Parameter = str_split(Parameter, "_", simplify = TRUE)[,1]) %>%
  # Split out each covariate
  separate_rows(`Covariate List`, sep =  ", ") %>%
  # Convert to uppercase
  mutate(`Covariate List` = toupper(`Covariate List`)) %>%
  # Convert 'none' to NA
  separate(`Covariate List`, c("Category", "Scale", "CovObsType", "Complexity"), sep = "_") %>%
  separate(`Category`, c("CovCategory", "EnvCategory"), sep = "-") %>%
  select(`Review ID`, Title, `Total covariates`, Parameter, CovCategory, EnvCategory, Scale, 
         CovObsType, Complexity)

@@

## Modelling
This contains information about the model selection and evaluation process
```{r Process modelling, echo = FALSE}
modelling_sheet <- spreadsheet[["Modelling"]] %>%
  filter(Status == "Complete") %>%
  mutate(`Prefiltering category` = tolower(`Prefiltering category`),
         `Selection category` = tolower(`Selection category`)) %>%
  select(`Review ID`, Title, Bayesian, `Prefiltering category`, `Selection category`, `Evaluation method`,
         `Multi-species`, Connectivity)

@@

```{r Save output, echo = FALSE}
reviewResults <- list(Study = study_sheet,
                      Objectives = objectives_sheet,
                      Taxa = taxa_location_sheet,
                      Survey = survey_sheet,
                      Covariates = covariates_sheet,
                      Modelling = modelling_sheet)
